---
title: "Mini-Project 2"
author: "Chujun He, Xuxi Ding, Guangzhi Sun "
date: "October 31, 2017"
output: html_document
---

```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(scales)
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

###Data Wrangling

By observing the contribution data frame, we find that most of transaction dates are in 2011 and 2012. Thus, we suppose that the contributions will be most relevant to the election in 2012. 

```{r,message=FALSE}
 #select certain columns, only need the information we want
 #only keep candidates whose election year is 2012
 #only examine the Republican party and Democratic party
 candidate_2012 <- candidates %>%
   select(cand_id,cand_name,cand_party_affiliation,cand_election_yr,cand_state) %>%
   filter(cand_election_yr == '2012', cand_party_affiliation == 'REP'|cand_party_affiliation == 'DEM')
 
# function that calculate how much certain type transactions did DEM/REP receive
  party_type_2012 <- function(input_type_1,input_2){
   contributions %>% 
     filter(transaction_type == input_type_1 | transaction_type == input_2) %>%   #only consider certain type transaction, which is directly used to support/oppose the candidates
      select(name,transaction_amt,cand_id,state) %>% #select certain columns, only need the information we want
      group_by(cand_id) %>% 
      summarize(total_trans = sum(transaction_amt)) %>%  #calculate total transaction amount given to a candidate
      right_join(candidate_2012, by ='cand_id') %>%  #join two data frames 
      filter(!is.na(total_trans)) %>%  #filter out NA values in total_trans
      group_by(cand_party_affiliation)%>% 
      summarize(party_trans = sum(total_trans))  #calculate the total amount of transactions to a party
  }

# 24E type transactions given to DEM/REP
party_24E_2012 <- party_type_2012('24E','')
# 24A type transactions given to DEM/REP
party_24A_2012 <- party_type_2012('24A','')
# Sum of 24E and 24A type transactions given to DEM/REP
party_24AE_2012 <- party_type_2012('24A','24E')
```

### Plot the graph
```{r,message=FALSE}
#function that plot certain type transactions given to support elections of two parties from 2011 to 2012
my_plot <- function(input_data){
   ggplot(input_data, aes(x=cand_party_affiliation, y= party_trans, fill = cand_party_affiliation)) +
    geom_bar(stat = "identity", width = 0.3) +
    scale_x_discrete(name='Party', label = c('Democratic', 'Republican'))+
    scale_y_continuous(name='Transactions', labels = scales::comma) +
    scale_fill_manual(values = c('blue', 'red'))
}

my_24E_plot <- my_plot(party_24E_2012) + ggtitle('Total Amount of Received 24E Transactions from 2011 to 2012')
my_24E_plot
```



(24E type transactions:  Independent expenditure advocating election of candidate)

```{r, message=FALSE}
my_24A_plot <- my_plot(party_24A_2012) + ggtitle('Total Amount of Received 24A Transactions from 2011 to 2012')
my_24A_plot
```


(24A type transactions: Independent expenditure opposing election of candidate)


```{r,message=FALSE}
my_24AE_plot <- my_plot(party_24AE_2012) + ggtitle('Total Amount of Received 24A/24E Transactions from 2011 to 2012')
my_24AE_plot
```

### Data Wrangling
```{r,message=FALSE}
#select certain columns, only need the information we want
#only donsider 24E/24A type transaction
#calculate total transaction amount given to a candidate
  contributions_24ae <- contributions %>%
    filter(transaction_type == "24E" | transaction_type == '24A') %>%
    select(name,transaction_amt,cand_id,state) %>% 
    group_by(cand_id) %>% 
    summarize(total_trans = sum(transaction_amt))
  
  
#join two data frames 
can_contri_2012 <- candidate_2012 %>% left_join(contributions_24ae, by = 'cand_id')
#filter out NA values in total_trans
can_contri_2012 <- can_contri_2012 %>% filter(!is.na(total_trans))

# function that calculates how much transactions did a party receive in each state
# arrange the data by the amount of transactions
 party_state_trans <- function(input_data,input_party){
   input_data %>% 
     filter(cand_party_affiliation == input_party) %>%
     group_by(cand_state) %>% 
     summarize(cand_tran = sum(total_trans)) %>% 
     filter(!is.na(cand_state)) %>%
     arrange(desc(cand_tran))
   }
 
# Republican received how much transactions in each state
rep_state_tran <- party_state_trans(can_contri_2012,'REP')
#rename the column
rep_state_tran$cand_tran_rep <- rep_state_tran$cand_tran
rep_state_tran$cand_tran <- NULL

# Demoncratic received how much transactions in each state
dem_state_tran <- party_state_trans(can_contri_2012,'DEM')
#rename the column
dem_state_tran$cand_tran_dem <- dem_state_tran$cand_tran
dem_state_tran$cand_tran <- NULL
 
two_states_tran <- rep_state_tran %>% full_join(dem_state_tran, by = 'cand_state')

#Wrangle the dataset to draw graph
Party_state <- two_states_tran %>%
  gather(key = "Party", value = "contributions",-cand_state) %>%
  group_by(cand_state)

# Look at the states where the contributions are greater than or equal to 1000000 (relatively large)
G_Party_state <- Party_state %>%
  filter(contributions >= 1000000) %>%
  filter(cand_state != '')
```

### Plot Data
```{r,message=FALSE}
#Draw the graph
my_plot <- ggplot(G_Party_state, aes(x = Party, y = contributions, fill = Party)) +
  geom_bar(stat = 'identity') + 
  scale_fill_manual(values = c('blue', 'red')) +
  scale_y_continuous(name='Transactions', labels = scales::comma) +
  scale_x_discrete(name='Party', label = c('DEM', 'REP'))+
  theme(axis.text.x = element_text(angle = 70,hjust = 1))+
  ggtitle('The Amount of Large Received Transactions(24A/E) by DEM and REP')+
  facet_wrap(~cand_state, ncol = 10)

my_plot

```

###Analysis and BG Info:
**Political candidates usually declare that money does not matter, yet the opposite is true. The influence of money in politics is, in fact, huge.**

In this mini-project, we find that a party or a candidate that out-fundraised its opponent is more likely to win campaigns. In other words, a winning party or candidate outspent its opponents.
(Theoretically, there are a variety of reasons for why the better-financed candidates are more likely to win. In many cases, the candidates who win the most votes do so-base on the same electability, popularity and qualifications that make them the best at fundraising.)

Where does the money come from? The money often does not come from candidate's net worth. A large amount of money comes from political action committees. *A political action committee (PAC) is a group formed for the purpose of contributing money to the campaigns of federal politicians.* It can spend money on either electing or on defeating a candidate. In the United States, a political action committee needs to be registered with the Federal Election Commission and spend more than $1,000 for the purpose of influencing a federal election.

We plotted a graph about the amount of large received transaction to support the democratic party and the Republican party. The graph is categorized in terms of the state. During the 2012 election, Barack Obama's home state was Illinois, Mitt Romney's home state was Massachusetts. **We find that Obama and Romney received a higher donation in their home state. $28,347,736 in Illinois and $70,528,037 in Massachusetts respectively.** The Republican party won the Democratic party at home state twice as much.

We also plotted a graph illustrating the independent expenditure opposing (24A) /advocating(24E) election of a candidate. During 2012 presidential campaign, independent groups have spent millions on television advertisements and other efforts asking voters to elect or defeat candidates. Super PACs, or independent-expenditure only committees, raised unlimited money from individuals, corporations and labor unions. Independent-expenditure only committees make independent expenditures in federal races, such as running ads or sending mail or communicating in other ways with messages that specifically advocate the election or defeat of a specific candidate. Though the Democratic party received less donation in 24E type transactions than the Republican party, it received more 24A type transactions. And the total amount of 24A and 24E transactions given to the Democratic party is more than that of the Republican party. Thus, we can say that during the election, super PACs supporting the Democratic party outspends super PACs supporting the Republican party on television advertisements and other efforts asking voters to defeat candidates. **This fits in our finding that better-financed candidate or parties are more likely to win campaigns.**