---
title: "Mini-Project 2"
author: "Chujun He, Xuxi Ding "
date: "October 31, 2017"
output: html_document
---

```{r, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(scales)
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

###Data Wrangling

By observing the contribution data frame, we find that most of transaction dates are in 2011 and 2012. Thus, we suppose that the contributions will be most relevant to the election in 2012. 

```{r,message=FALSE}
 #select certain columns, only need the information we want
 #only keep candidates whose election year is 2012
 #only examine the Republican party and Democratic party
 candidate_2012 <- candidates %>%
   select(cand_id,cand_name,cand_party_affiliation,cand_election_yr,cand_state) %>%
   filter(cand_election_yr == '2012', cand_party_affiliation == 'REP'|cand_party_affiliation == 'DEM')
 
# function that calculate how much certain type transactions did DEM/REP receive
  party_type_2012 <- function(input_type_1,input_2){
   contributions %>% 
     filter(transaction_type == input_type_1 | transaction_type == input_2) %>%   #only consider certain type transaction, which is directly used to support/oppose the candidates
      select(name,transaction_amt,cand_id,state) %>% #select certain columns, only need the information we want
      group_by(cand_id) %>% 
      summarize(total_trans = sum(transaction_amt)) %>%  #calculate total transaction amount given to a candidate
      right_join(candidate_2012, by ='cand_id') %>%  #join two data frames 
      filter(!is.na(total_trans)) %>%  #filter out NA values in total_trans
      group_by(cand_party_affiliation)%>% 
      summarize(party_trans = sum(total_trans))  #calculate the total amount of transactions to a party
  }

# 24E type transactions given to DEM/REP
party_24E_2012 <- party_type_2012('24E','')
# 24A type transactions given to DEM/REP
party_24A_2012 <- party_type_2012('24A','')
# Sum of 24E and 24A type transactions given to DEM/REP
party_24AE_2012 <- party_type_2012('24A','24E')
```

### Plot the graph
```{r,message=FALSE}
#function that plot certain type transactions given to support elections of two parties from 2011 to 2012
my_plot <- function(input_data){
   ggplot(input_data, aes(x=cand_party_affiliation, y= party_trans, fill = cand_party_affiliation)) +
    geom_bar(stat = "identity", width = 0.3) +
    scale_x_discrete(name='Party', label = c('Democratic', 'Republican'))+
    scale_y_continuous(name='Transactions', labels = scales::comma) +
    scale_fill_manual(values = c('blue', 'red'))
}

my_24E_plot <- my_plot(party_24E_2012) + ggtitle('Total Amount of Received 24E Transactions from 2011 to 2012')
my_24E_plot
```

Firstly, the figure of independent expenditure of advocating election of candidate shows that Democratic party reveived less donations than republic party sigificantly. As we assumed, the more donation a party receives, the more likelihood it will win. Since the Obama won the election, we naturally need to draw a big picture to include donation from other channels. 

```{r, message=FALSE}
my_24A_plot <- my_plot(party_24A_2012) + ggtitle('Total Amount of Received 24A Transactions from 2011 to 2012')
my_24A_plot
```

secondly, the figure of independent expenditure of oppising election of candidate shows that Democratic party collected much higher level of donations than republic party did. This result coincides with our assumption.

```{r,message=FALSE}
my_24AE_plot <- my_plot(party_24AE_2012) + ggtitle('Total Amount of Received 24A/24E Transactions from 2011 to 2012')
my_24AE_plot
```
Adding those two parts - independent expenditure of oppising and advocating election of candidate - together, democratic pary outweights republic party in terms of fundrising. we confirm our assumption that the more donation a party receives, the more likelihood it will win.

###Our analysis and bg information:
**Democratic recevied more transactions than Republic in 2012.**
In the United States presidential election of 2012, the Democratic nominee, incumbent President Barack Obama was elected to a second term, defeating the Republican nominee, former Governor of Massachusetts Mitt Romney. Intutively, the more transactions given to support elections of a party, the higher chance that the donation can be used to support the party to win the election. However, the nominee of Democratic party, Barack Obama, won the election with less transactions.


### Data Wrangling
```{r,message=FALSE}
#select certain columns, only need the information we want
#only donsider 24E/24A type transaction
#calculate total transaction amount given to a candidate
  contributions_24ae <- contributions %>%
    filter(transaction_type == "24E" | transaction_type == '24A') %>%
    select(name,transaction_amt,cand_id,state) %>% 
    group_by(cand_id) %>% 
    summarize(total_trans = sum(transaction_amt))
  
  
#join two data frames 
can_contri_2012 <- candidate_2012 %>% left_join(contributions_24ae, by = 'cand_id')
#filter out NA values in total_trans
can_contri_2012 <- can_contri_2012 %>% filter(!is.na(total_trans))

# function that calculates how much transactions did a party receive in each state
# arrange the data by the amount of transactions
 party_state_trans <- function(input_data,input_party){
   input_data %>% 
     filter(cand_party_affiliation == input_party) %>%
     group_by(cand_state) %>% 
     summarize(cand_tran = sum(total_trans)) %>% 
     filter(!is.na(cand_state)) %>%
     arrange(desc(cand_tran))
   }
 
# Republican received how much transactions in each state
rep_state_tran <- party_state_trans(can_contri_2012,'REP')
#rename the column
rep_state_tran$cand_tran_rep <- rep_state_tran$cand_tran
rep_state_tran$cand_tran <- NULL

# Demoncratic received how much transactions in each state
dem_state_tran <- party_state_trans(can_contri_2012,'DEM')
#rename the column
dem_state_tran$cand_tran_dem <- dem_state_tran$cand_tran
dem_state_tran$cand_tran <- NULL
 
two_states_tran <- rep_state_tran %>% full_join(dem_state_tran, by = 'cand_state')

#Wrangle the dataset to draw graph
Party_state <- two_states_tran %>%
  gather(key = "Party", value = "contributions",-cand_state) %>%
  group_by(cand_state)

# Look at the states where the contributions are greater than or equal to 1000000 (relatively large)
G_Party_state <- Party_state %>%
  filter(contributions >= 1000000) %>%
  filter(cand_state != '')
```

### Plot Data
```{r,message=FALSE}
#Draw the graph
my_plot <- ggplot(G_Party_state, aes(x = Party, y = contributions, fill = Party)) +
  geom_bar(stat = 'identity') + 
  scale_fill_manual(values = c('blue', 'red')) +
  scale_y_continuous(name='Transactions', labels = scales::comma) +
  scale_x_discrete(name='Party', label = c('DEM', 'REP'))+
  theme(axis.text.x = element_text(angle = 70,hjust = 1))+
  ggtitle('The Amount of Large Received Transactions(24A/E) by DEM and REP')+
  facet_wrap(~cand_state, ncol = 10)

my_plot

```

###Analysis and BG Info:
**Candidates usually received large amount of money in their home states.**
Barack Obama's(Deomoncratic Candidate) home state was Illinois, while Mitt Romney's(Republican Candidate) home state was Massachusetts. By observing the candidate data frame, we find that Obama and Romney received the highest donation in their home state: Illinois (28347736) and Massachusetts (70528037) respectivcely; and that Republic party won deomate at home state as twice much as democartic party did. This may help to explain that democratic party had less amount of donation in general. We consider that is there any other factors contribute to winning election. Thus, we suppose that the other factors will may be relevant to the election in 2012.

The United States presidential election of 2012 broke new records in fundraising. Through grassroots campaign contributions, online donations, and Super PACs, Obama won significant higher amount of small individual contributions than Romney did. 

#Note: The exact fact is: Obama won more donations than Romney ($722,393,592	versus $449,886,513), but our data shows a different result. We can discuss it more tomorrow. 
#http://www.opensecrets.org/pres12/




